---
title: "some troubleshooting"
author: 'An Bui'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries-data, message = FALSE}
# libraries
library(tidyverse)
library(vegan)
library(here)

# data
# algal  communities
PointLoma <- read_csv(here::here("AB-troubleshooting", "PL_div.csv")) %>% 
  column_to_rownames("Site")

# environmental variables
pl_meta <- read_csv(here::here("AB-troubleshooting", "PL_div_meta.csv"))

# set up a "metadata" frame - will be useful for plotting later!
site_type <- pl_meta %>% 
  # selects the named columns
  select(Site, Region, Plot, Treatment, Season, Year)
```

#### NMDS

Non-metric Multidimensional Scaling  

Ok, so this is kinda tricky because it turns out that some of your sites had 0 observations. So, not an issue with your species! I filtered out sites with 0 observations by summing all observations for each site, then filtering out any row with a sum of observations greater than 0 (essentially only keeping sites with some algae counted).  

```{r algae-NMDS}
PointLoma1 <- PointLoma %>% 
  # filter out sites with 0 observations
  mutate(sum = rowSums(across(where(is.numeric)))) %>% 
  filter(sum > 0) %>% 
  select(-sum) 
```

Then, I pulled a vector of the sites that were kept (seems like there are 409 of them). This is going to be useful later, when the `meta` dataframe has to be filtered for the sites that are included in the NMDS.
```{r}
# pull row names from PointLoma1 (which will tell you which sites DID have observations)
kept_sites <- PointLoma1 %>% 
  # turn row names back into a column
  rownames_to_column("site") %>% 
  # pull the site column as a vector
  pull(site)

kept_sites
```

Now the NMDS! Should be working.
```{r}
alg_NMDS <- metaMDS(PointLoma1)

alg_NMDS
```

Here's where I filter the `meta` data frame for all sites in `kept_sites`:
```{r}
kept_pl_meta <- pl_meta %>% 
  filter(Site %in% kept_sites)
```

And now a plot:
```{r NMDS-plot}
# extract outputs of nMDS into data frame
nmds_df <- as_tibble(alg_NMDS$points) %>% 
  # binding it with metadata
  bind_cols(kept_pl_meta, .)

nmds_plot <- ggplot(nmds_df, aes(x = MDS1, y = MDS2, color = Treatment, shape = Treatment)) +
  geom_point(size = 3, alpha = 0.8) +
  stat_ellipse()
nmds_plot
```

```{r}
#PERMANOVA
pl.div <- adonis2(PointLoma1 ~Treatment*Season, data = kept_pl_meta,
                  permutations = 999, method = "bray")
pl.div
```

```{r}
#dispersion
pl.dist <- vegdist(PointLoma1,
                   method = "bray")

dispersion <- betadisper(pl.dist, group=kept_pl_meta$Season)
permutest(dispersion)
plot(dispersion, hull=FALSE, ellipse=TRUE) ##sd ellipse

```

